\documentclass[12pt,letterpaper]{article}
\usepackage{mla}
\usepackage{ifpdf}
\usepackage{setspace}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\doublespacing
\begin{document}
\begin{mla}{Jonah}{Eadie}{Ms. Smith}{English 2 HP, Period 2}{2 February 2015}
{On William Golding's Use of Jack's Varying Position of Power in \textit{The 
Lord of the Flies}}

It has often been said that those who seek positions of power are those who are 
least fit to yield it. In \textit{The Lord of the Flies}, William Golding 
attempts to portray this central idea of unfit and corrupt leadership, with 
much of the novel's drama being found in the power struggle between the piece's 
central characters, Ralph and Jack. Jack's attempts to usurp leadership 
provides for the novel's most biting social commentary: Jack, being one who has 
always found himself in positions of control, naturally expects that he will 
reign as the ruler of the castaway boys; and as the story progresses, Jack 
manages, through ruthless maneuvering, to shift the balance of power away from 
Ralph and towards himself. This overarching narrative of unscrupulous 
megalomaniacs clawing their way to positions of power acts as a commentary on 
the modern political climate.

In the beginning of the novel, Jack naturally sees himself as the most fit 
ruler among the boys given his previous position of leadership in the choir, 
this being indicative of the “old-guard” power trope that mirrors the power 
struggles one finds in modern political life. When one is first introduced to 
Jack, one is given the image of a “natural leader”; he is found leading the 
choir boys, his most vehement supporters, to Ralph and Piggy's tribunal. When 
deciding upon leadership, Jack is overtly upset by the group's decision to cast 
Ralph as their leader, and just barely manages to contain his frustration over 
the decision. This inclusion of Ralph's anger over the boys’ vote by Golding 
neatly sets the tone of the two's power struggle found throughout the piece as 
its central drama, and gives the reader insight into Jack's privileged and 
self-entitled placement of himself within his world. In effect, Jack acts as a 
symbol for the “old-guard” within the power structures and political system of 
the modern world. Jack, fully expecting power and respect simply because of his 
past experiences with power, represents aristocracy. The power he gains is not 
derived from merit or public opinion, but rather from past history of power and 
his own cunning. This directly plays into the author's intended message of rule 
by the unjust.

Jack, through the course of the novel, uses his cunning and the inherited 
power granted to him through the choirboys to unseat Ralph, the elected leader 
of the boys, and places himself at the top of the island's sociopolitical 
food-chain, providing an analogue to modern political dealings one finds in the 
politics of today. Additionally, while Jack initially acts as a friend and ally 
to Ralph, the reader soon finds that this friendship ends when Jack realizes 
that Ralph will not be as easy to manipulate and to control as he finds the 
other boys. Thusly, Ralph stands his ground on the issue of the signal fire, a 
move that was not anticipated by Jack. This mirrors the political bullying one 
finds in modern political systems, with unjust actors, through immoral and 
amoral dealings, vying for positions of power that they otherwise would not be 
able to obtain.

It can be seen, then, that Golding used the character of Jack as an analogue 
for corrupt political actors.


\end{mla}
\end{document}
